---
title: "Activity 8"
author: "Bruce Buurstra"
output: github_document
---

## Data and packages

Load the entirety of the `{tidyverse}`.
You will do a little work with dates in this activity so also load `{lubridate}`.
Be sure to avoid printing out any unnecessary information and give the code chunk a meaningful name.

```{r load packages, message=FALSE}
library(tidyverse)
library(lubridate)
```


Cheatsheets that you might want to add to your collection:

- [`{stringr}`](https://stringr.tidyverse.org/index.html)
- [`{forcats}`](https://forcats.tidyverse.org/)

Using `here::here`, upload the `billboard_songs.txt` file that is saved in your `data/` folder.
Notice that this file is a tab-delimited file that is stored as a `.txt` file.
Therefore, you will need to use either `read_delim` with a `delim = ...` argument or (better) `read_tsv`.
Assign the file to a meaningful object name, be sure to avoid printing any unnecessary information, and give the code chunk a meaningful name.


```{r load data set}
billboard_songs <- read_tsv(here::here("data", "billboard_songs.txt") )
```

These data include information on song popularity.
In the US, the Billboard Hot 100 is a list that comes out every week, showing the 100 most played songs that week.
More information about the creation of this dataset, as well as some analyses by the author, can be found here: [https://mikekling.com/analyzing-the-billboard-hot-100/](https://mikekling.com/analyzing-the-billboard-hot-100/).
The dataset you are provided is a limited version of the full data, containing:

- `title`
- `artist`
- `overall peak`: The highest rank the song ever reached (1 is the best)
- `weeks on chart`: The number of weeks the song was on the chart
- `chart date`: The latest date the song appeared on the Billboard Hot 100

This is a long dataset (34,605 observations)!

## Analysis

### Cleaning date

When you look at ``chart date`` you'll notice that it was read in as a numeric (`<dbl>`) variable type.
What does the format of the date appear to be in?

**Response**:ymd


In this activity, we will be using `str_...` functions to split or find patterns in strings of information.
Do the following with your dataset:

- *Create* a variable called `date` that treats ``chart date`` *as a character* variable type,
- Use `stringr::str_sub` on `date` to create three new columns: `year`, `month`, `day`
- Replace/overwrite the variable `date` with an updated format using `lubridate::make_date`.

Once you have verified that this worked, overwrite your data object.

```{r creating dates in data}
billboard_songs <- billboard_songs %>%
   mutate(date = as.character(`chart date`),
          year = str_sub(date, 1, 4), 
          month = str_sub(date, 5, 6), 
          day = str_sub(date, 7, 8), 
          date = make_date(year, month, day))
```


#### Analyzing using dates

What 10 songs spent the longest on the charts?
Give only the title, artists, and weeks.

```{r top10songs}
billboard_songs%>%
  arrange(desc(`weeks on chart`))%>%
  select(title, artist, `weeks on chart`)%>%
  head(10)
```


**Response**:

What date did the oldest song(s) in this dataset leave the charts?
Give only the *distinct* dates.

```{r}
billboard_songs %>%
  arrange(date)%>%
  #mutate(date_left = date + weeks(`weeks on chart`))%>%
  distinct(date)
```

**Response**:


### Artists who work together

Before we get to far, let's create common definitions:

In the string below, Nicki Minaj and Young Thug are considered to be **featured**.

```
RAE SREMMURD featuring NICKI MINAJ & YOUNG THUG
```

In the string below, Jessie J and Ariana Grande and Nicki Minaj are all considered to have **collaborated** on the song.

```
JESSIE J, ARIANA GRANDE & NICKI MINAJ
```

Which artist has been **featured** on the most Billboard charting songs?

First, create a vector object called `artist_credentials` that takes your dataset and *then* `pull`s only the `artist` column.
The `pull` function is like `select` except that it outputs a vector rather than a data frame (tibble).
Using this function will help us avoid many "`coercing`" warnings - I find it easier to work with vectors of strings.

```{r pull artist column}
artist_credentials <- billboard_songs %>%
  pull(artist)
```


Using the `artist_credentials` vector, do the following. I recommend that you do one bullet at time and verify that it worked.

- Use `str_subset` to pick only the observations that only include `"featuring"`,*then*
- Use `str_replace_all` to remove everything before and including "featuring" for each of these observations (i.e., replace it with `""`), *then*
- Use `str_split` to separate entries with multiple featured artist - these could include `,`, `&`, ` or ` (note the whitespace), or ` and ` (note the whitespace) (this should produce a list), *then*
- Use `unlist` to lengthen all featured artists into one vector (compare the output here to the output from the previous bullet point), *then*
- Use `str_trim` to remove whitespace from `"both"` sides of each entry.
?str_replace_all
When you verify that this works, save this information into an object called `feat_artist`.

```{r}
str1 <- str_subset(artist_credentials, "featuring")
str2 <- str_replace_all(str1, ".{1,}featuring", "")
str3 <- str_split(str2, ",")
str4 <- str_split(str3, "&")
str5 <- str_split(str4, " or ")
str6 <- str_split(str5, " and ")
str7 <- unlist(str6)
feat_artist <- str7
```

Now you can use this information to answer the question above.

```{r}
featured_artists <- as.data.frame(feat_artist)
featured_artists %>%
  group_by(feat_artist)%>%
  summarise(count = n())%>%
  arrange(desc(count))
```



**Response**:
Lil Wayne

Which artist has **collaborated** on the most Billboard charting songs?

Using the `artist_credentials` vector you created before, do the following:

- Use `str_replace_all` to remove everything after and including "featuring" for each of these observations (i.e., replace it with `""`), *then*
- Use `str_extract_all` to separate entries by `,` or `&`, *then*
- Use `unlist` to lengthen all collaborating artists into one vector, *then*
- Use `str_trim` to remove whitespace from `"both"` sides of each entry.

When you verify that this works, save this information into an object called `collab_artist`.

```{r}
str1 <- str_replace_all(artist_credentials, " featuring.{1,}", "")
str2 <- str_split(str1, ",")
str3 <- str_split(str2, "&")
str4 <- unlist(str3)
str5 <- str_trim(str4, side = "both")
coll_artist <- str5
```


```{r}
coll_artists <- as.data.frame(coll_artist)
coll_artists %>%
  group_by(coll_artist)%>%
  summarise(count = n())%>%
  arrange(desc(count))
```


Now you can use this information to answer the question above.

Glee Cast


#### Challenge 1

What songs could have been played at your 16th birthday party?
That is, which songs that eventually peaked at #1  **entered** the charts within a couple months (before or after) your 16th birthday?
Give only the song title, artist, and date of chart entry.

Note that this might not be possible depending on when your 16th birthday occurred (i.e., if you turned 16 after March 7, 2015 - I need to find more recent songs :/).





**Response**:



## Challenge 2: Data Exploration

This is somewhat unrelated to the steps you completed in the rest of this activity, but you have added experience with new tools that you may find useful.

Create some visual representation or summary table of these data.
I would encourage you to explore how you can use `{forcats}` to help make your visual/table more human-readable/effective for telling the story of your exploration.




## Attribution

Parts of this Activity are based on a lab from [Dr. Kelly Bodwin's](https://www.kelly-bodwin.com/) STAT 331 course.